Hydras \& Co.

### 2022-03-17 doc/part-hydras.tex
corrections chapter 4

It is natural to try to prove by structural induction over \texttt{T1} 
that every term in normal form is accessible through \texttt{LT}.

Unfortunately, it won't work. Let us consider some well-formed term
 $\alpha=\texttt{cons $\beta\;n\;\gamma$}$, and assume that \(\beta\) and \(\gamma\) are accessible
 through \texttt{LT}. For proving the accessibility of $\alpha$, we have to consider
any well formed term \(\delta\) such that \(\delta<\alpha\). 
But nothing guarantees that \(\delta\)  is strictly  less than \(\beta\) nor \(\gamma\), and we cannot use the induction hypotheses on   \(\beta\) nor \(\gamma\).

@@

The problem comes from the hypothesis \texttt{Hdelta}. It does not prevent  \(\delta\) to be bigger that \(\beta\) or
\(\gamma\);
for instance \(\delta\) may be of the form
\texttt{cons $\beta$ $p'$  $\gamma'$},
where    \(p' < n\).
Thus, the induction hypotheses \texttt{IHbeta} and \texttt{IHgamma}  are useless for finishing our proof.

@@

  Instead of trying to prove directly that any ordinal term \(\alpha\) in normal form is accessible
through \texttt{LT}, we propose to show first that any well formed 
term of the form \(\omega^\alpha\times(n+1)+\beta\) is accessible (which is a stronger result).

@@

As a corollary, the order \texttt{Lt} on type \texttt{Ord} is well-founded too.

\inputsnippets{E0/E0LtWf}.

@@

We build an instance of \texttt{ON}, and prove its correctness w.r.t. Schutte's model.

@@

@@

  \pagebreak 
  \vspace{4pt}
    \emph{From Module~\href{../theories/html/hydras.Epsilon0.Hessenberg.html\#oplus}{Epsilon0.Hessenberg}}

@@

    \subsubsection{Monotonicity}

@@

@@

The proof is as follows:
\begin{enumerate}
\item Since $\alpha<\epsilon_0$, consider the least $i$ such that $\alpha$ is strictly less than the omega-tower of height $i$.
\item
  \begin{itemize}
  \item If $i=0$, then the result is trivial (because $\alpha=0$)
 \item  Otherwise let $i=j+1$; 
          $\alpha$ is greater than or equal to the omega-tower of height $j$.
         By monotonicity,  $\phi_0(\alpha)$ is greater than or equal to 
        the omega-tower of height $j+1$, thus strictly greater than $\alpha$
  \end{itemize}
 \end{enumerate}

>>>>

It is natural to try to prove by structural induction over \texttt{T1} 
that every term in normal form is \texttt{LT}-accessible.

Unfortunately, it won't work. Let us consider some well-formed term
 $\alpha=\texttt{cons $\beta\;n\;\gamma$}$, and assume that \(\beta\) and \(\gamma\) are \texttt{LT}-accessible.
In order to prove the accessibility of $\alpha$, we have to consider
any well formed term \(\delta\) such that \(\delta<\alpha\). 
But nothing guarantees that \(\delta\)  is strictly  less than \(\beta\) nor \(\gamma\), and we cannot use the induction hypotheses on   \(\beta\) nor \(\gamma\).

@@

The problem comes from the too weak hypothesis \texttt{Hdelta}. It does not prevent  \(\delta\) to be bigger that \(\beta\) or
\(\gamma\);
for instance \(\delta\) may be of the form
\texttt{cons $\beta$ $p'$  $\gamma'$},
where    \(p' < n\).
Thus, the induction hypotheses \texttt{IHbeta} and \texttt{IHgamma}  are useless for finishing our proof.

@@

  Instead of trying to prove directly that any ordinal term \(\alpha\) in normal form is accessible
through \texttt{LT}, we propose to consider the following (stronger) predicate:

@@

As a corollary, the order \texttt{Lt} on type \texttt{E0} is well-founded too.

\inputsnippets{E0/E0LtWf}

@@

We build an instance of \texttt{ON}, and prove its correctness w.r.t. Schutte's model (see Chapter~\ref{chap:schutte}).

@@

    \subsection{Related work}
    The article~\cite{Manoury2016}  defines another representation of ordinals below $\omega^\omega$ based on lists of natural numbers.

@@

%  \pagebreak 
  \vspace{4pt}
    \emph{From Module~\href{../theories/html/hydras.Epsilon0.Hessenberg.html\#oplus}{Epsilon0.Hessenberg}}

@@

    \subsubsection{Monotony}

@@

    \paragraph*{\gaiasign} The module
      \href{../theories/html/gaia_hydras.~GHessenberg.html}{gaia\_hydras.GHessenberg} defines a version of Hessenberg's sum
      compatible with \gaia's type \texttt{T1}.

@@

The proof is as follows:
\begin{enumerate}
\item Since $\alpha<\epsilon_0$, consider the least $i$ such that $\alpha$ is strictly less than the omega-tower of height $i$.
\item
  \begin{itemize}
  \item If $i=0$, then the result is trivial (because $\alpha=0$)
 \item  Otherwise let $i=j+1$; 
          $\alpha$ is greater than or equal to the omega-tower of height $j$.
         By monotony,  $\phi_0(\alpha)$ is greater than or equal to 
        the omega-tower of height $j+1$, thus strictly greater than $\alpha$
  \end{itemize}
 \end{enumerate}
